---
title: "Topic 3: Crime Rate vs Crime Count"
output: html_notebook
---



```{r}
pop <- read_csv("Data/Census Population/Data_UNIT_URESPOP.csv") %>% slice(3:57) %>% 
  select(2,3,6,7, 8) %>%
  janitor::clean_names() %>%
  rename(pop_density = f2383, 
         pop_count_wrk = f2384, 
         pop_count_res = f323339) %>% 
  mutate_at(c('pop_density', 'pop_count_wrk',
              'pop_count_res'), as.numeric)
```



```{r}
surrey_lsoa <- left_join(surrey_lsoa, pop, by = c("code"="geo_code"))

```



```{r}
surrey_lsoa <- surrey_lsoa %>% 
  mutate(crime_rate = (count/pop_count_wrk)*1000)

```



##ggplot

```{r}
ggplot() + 
  annotation_map_tile() + 
  geom_sf(data = surrey_lsoa, aes(fill = crime_rate), alpha = 0.5) + 
  scale_fill_gradient2(name ="Crime Rate")
```



##tmaps 

```{r}
tm_shape(surrey_lsoa) + 
  tm_fill("crime_rate", style = "quantile") + 
  tm_borders(alpha = 0.3)
```




## Cartograms and ggplot


```{r}
library(cartogram)

#In our data set we have a variable “pop_count” which refers to the total number of people in our LSOA 
cart <- cartogram_cont(surrey_lsoa, weight = "pop_count_wrk")   

## simple plot 
ggplot(cart) + 
  geom_sf()


## fill with our count variable 
ggplot(cart) + 
  geom_sf(aes(fill = pop_count_wrk))


## add in some aesthetics 
ggplot(cart) + 
  geom_sf(aes(fill = pop_count_wrk), 
          color = "gray50", 
          linetype = 1, 
          lwd = 0.35) + 
  scale_fill_gradientn(colours = heat.colors(n =10, 
                                            alpha = 0.5, 
                                            rev = TRUE)) + 
  theme_gray() + 
  labs(title = "Surrey Heath: Population by LSOA", 
       subtitle = "August 2020")
```







## Activity ? 

We have mapped the variable pop_count_wrk, now lets do the same with the variable pop_count_res

Steps: 
- 1) First calculate the crime rate 
- 2) Plot using ggplot 
- 3) Plot using tmap 
- 4) Plot both maps together using tmap_arrange 
- 5) Plot a cartogram of residential population

Is there a difference between the crime rate when using workday population compared to residential population? Would we expect to see these trends




```{r}
# 1) First calculate the crime rate 

surrey_lsoa <- surrey_lsoa %>% 
  mutate(crime_rate2 = (count/pop_count_res)*1000)


#2) Plot using ggplot 

ggplot() + 
  annotation_map_tile() + 
  geom_sf(data = surrey_lsoa, aes(fill = crime_rate2), alpha = 0.5) + 
  scale_fill_gradient2(name ="Crime Rate")


#3) Plot using tmap 

tm_shape(surrey_lsoa) + 
  tm_fill("crime_rate", style = "quantile") + 
  tm_borders(alpha = 0.3)


#4) Compare the workday vs residential population 

e <- tm_shape(surrey_lsoa) + 
  tm_fill("crime_rate", style = "quantile", title = "Workday Pop") + 
  tm_borders(alpha = 0.3)

f <- tm_shape(surrey_lsoa) + 
  tm_fill("crime_rate2", style = "quantile", title = "Residential pop") + 
  tm_borders(alpha = 0.3)


#5) Plot both maps together using tmap_arrange 
tmap_arrange(e, f)



#6) 

ggplot(cart) + 
  geom_sf(aes(fill = pop_count_res), 
          color = "gray50", 
          linetype = 1, 
          lwd = 0.35) + 
  scale_fill_gradientn(colours = heat.colors(n =10, 
                                            alpha = 0.5, 
                                            rev = TRUE)) + 
  theme_gray() + 
  labs(title = "Surrey Heath: Population by LSOA", 
       subtitle = "August 2020")

```
 

