---
title: "Extra Topics and Next Steps"
output: html_notebook
---




### Binning data - maybe include, to complicaed to include I think

Binning, can be thought of as a two-dimensional histogram (shading of the bins take the heights of the bars). 

Need to convert the sf data.frame geometry column into a data.frame with separate x, y columns 

How do you separate the coordinates? 

https://github.com/r-spatial/sf/issues/231 

```{r}
sfc_as_cols <- function(x, names = c("x","y")) {
  stopifnot(inherits(x,"sf") && inherits(sf::st_geometry(x),"sfc_POINT"))
  ret <- sf::st_coordinates(x)
  ret <- tibble::as_tibble(ret)
  stopifnot(length(names) == raster::ncol(ret))
  x <- x[ , !names(x) %in% names]
  ret <- setNames(ret,names)
  dplyr::bind_cols(x,ret)
}


sf_seperate <- sfc_as_cols(sf, c("lng", "lat")) 


ggplot(sf_seperate, aes(lng, lat)) +   
  annotation_map_tile() +
  stat_binhex(bins = 30) +                                           
  scale_fill_gradientn(colours = c("white","red"), name = "Frequency")   


#hexagonal = stat_binhex() 
#rectangle = stat_bin2d()
#heat = stat_density2d()  

```




# Extra Topic: Interactive Maps; Leaflet

```{r}

## Subsetting for just ASB 
asb <- subset(crime01_19, crime_type == "Anti-social behaviour")

m <- leaflet(data = asb) %>%
  addProviderTiles("Stamen.Toner") %>% 
  addMarkers(lng=~longitude, lat=~latitude, popup=~as.character(location), label = ~as.character(location))
m

```






## Next Steps 

1. Explore Trends across different months (small multitudes)




